Left-linear rules in adhesive categories

### 2024-02-19 icalp/fusione.tex
typos in proof

\usepackage{hyperref}
\usepackage[disable,textsize=tiny]{todonotes}
%\usepackage{etex}
\usepackage[all]{xy}
\SelectTips{cm}{}
% derivations and labels
\usepackage{proof}
\newcommand{\lab}[1]{\ensuremath{\mathsf{{#1}}}}
\newcommand{\slab}[1]{\ensuremath{\scriptstyle{\mathsf{{#1}}}}}
\usepackage{wrapfig}

@@

  A \emph{left-linear rewriting system} is a pair $(\X, \R)$ where
  $\X$ is an $\mathcal{M}$-adhesive category and $\R$ a set of left
  $\mathcal{M}$-linear rules. The rewriting system $(\X, \R)$ is
  called $\mathcal{M}$-\emph{linear}\todo{P: giusto?}
  %
  \full{\emph{(consuming)}}
  %
  if every rule in $\R$ is
  $\mathcal{M}$-linear.
  \full{(consuming)}

@@

\begin{restatable}[No need of useless switchings]{theorem}{thNoNeed}
  \label{th:no-need}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If $\der{D} \shifteq \der{D}'$ then there is a switching sequence consisting only of inversions
  \begin{center}
    $\der{D} \equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n \equiv_a \der{D}'$
  \end{center}
\end{restatable}

@@

Still, we can identify a
canonical form for switching derivations: at each step one can apply a switch to the inversion with higher index.

\begin{restatable}[Canonical form]{corollary}{coCanonical}
  \label{co:canonical}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If
  $\der{D} \shifteq[\sigma]\der{D}'$ then there is a switching sequence
  \begin{center}
    $\der{D} \equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n \equiv_a\der{D}'$
  \end{center}
  where for each $i \in \interval{n}$ the transposition $\nu_i  = \transp{k}$ is such that 
  $k = \max \{ j \mid \transp{j} \in \inv{\nu_{i,n}}\}$.\todo{P: $\nu_{i,n} = \sigma \circ \nu_{1,i}$, giusto? forse e' piu' carino}
\end{restatable}

@@

There is another
useful property of  sequentially independent direct
derivations which is worth mentioning.\todo{A: va tolto weakly?}

@@

  By Lemma~\ref{lem:indep-global-left}(\ref{lem:indep-global-left:1})\todo{Check numbering} we can
  reorganise the first three switchings as

@@

  Since in $\der{D}_0$ steps in positions $i$ and $i+1$ are sequential
  independent, we can use
  Lemma~\ref{lem:indep-global-left}(\ref{lem:indep-global-left:1})\todo{Check
    numbering} and we have that the corresponding steps in positions
  $i+1$ and $i+2$ in $\der{D}_2''$ are still sequential
  independent. Hence, we can use the inductive hypothesis on
  $\der{D}_2''$ and deduce the existence of a switching sequence where
  $\transp{i+1}[i+2]$ can be switched first, i.e. we obtain (see
  Fig.~\ref{fi:switchB2}, third column)
  \begin{center}
    $\der{D}_2'' \shift{\transp{i+1}[i+2]}
    \der{D}_3''' \shift{\nu_4'''} \ldots \shift{\nu_n'''} \der{D}_n$
  \end{center}

  Putting things together, we obtain
  \begin{center}
    $\der{D}_0 \shift{\transp{i+1}[i+2]} \der{D}_1' \shift{\transp{i}}
    \der{D}_2'' \shift{\transp{i+1}[i+2]} \der{D}_3''' \shift{\nu_4'''}\ldots \shift{\nu_n'''} \der{D}_n$
  \end{center}
  Since we additionally know that in $\der{D}_0$ the steps $i$ and
  $i+1$ are sequentially independent, we can use
  Lemma~\ref{lem:indep-global-left}(\ref{lem:indep-global-left:2})\todo{Check numbering} to
  reorder the first three switchings, obtaining
  \begin{center}
    $\der{D}_0 \shift{\transp{i}} \der{D}_1' \shift{\transp{i+1}[i+2]}
    \der{D}_2'' \shift{\transp{i}} \der{D}_3''' \shift{\nu_4'''}\ldots \shift{\nu_n'''} \der{D}_n$
  \end{center}
  as desired.  The fact that the three switchings above lead to
  $\der{D}_3'''$ is a consequence of Lemma~\ref{lem:switch-confluence}.
\end{proof}

>>>>

\usepackage{hyperref}
\usepackage[%disable,%
textsize=tiny]{todonotes}
%\usepackage{etex}
\usepackage[all]{xy}
\SelectTips{cm}{}
% derivations and labels
\usepackage{proof}
\newcommand{\lab}[1]{\ensuremath{\mathsf{{#1}}}}
\newcommand{\slab}[1]{\ensuremath{\scriptstyle{\mathsf{{#1}}}}}
\usepackage{wrapfig}

@@

  A \emph{left-linear rewriting system} is a pair $(\X, \R)$ where
  $\X$ is an $\mathcal{M}$-adhesive category and $\R$ a set of left
  $\mathcal{M}$-linear rules. The rewriting system $(\X, \R)$ is
  called $\mathcal{M}$-\emph{linear}
  %
  \full{\emph{(consuming)}}%
  %
  if every rule in $\R$ is
  $\mathcal{M}$-linear.
  \full{(consuming)}

@@

\begin{restatable}[No need of useless switches]{theorem}{thNoNeed}
  \label{th:no-need}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If $\der{D} \shifteq \der{D}'$ then there is a switching sequence consisting only of inversions
  \begin{center}
    $\der{D} \equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n \equiv_a \der{D}'$
  \end{center}
\end{restatable}

@@

Still, we can identify a
canonical form for switching derivations: at each step one can apply a switch to the inversion with largest index.

\begin{restatable}[Canonical form]{corollary}{coCanonical}
  \label{co:canonical}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If
  $\der{D} \shifteq[\sigma]\der{D}'$ then there is a switching sequence
  \begin{center}
    $\der{D} \equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n \equiv_a\der{D}'$
  \end{center}
  where for each $i \in \interval{n}$ the transposition $\nu_i  = \transp{k}$ is such that 
  $k = \max \{ j \mid \transp{j} \in \inv{\nu_{i,n}}\}$.
\end{restatable}

@@

There is another
useful property of  sequentially independent direct
derivations which is worth mentioning.

@@

  By Lemma~\ref{lem:indep-global-left}(\ref{lem:indep-global-left:1})
  we can reorganise the first three switchings as

@@

  Since in $\der{D}_0$ steps in positions $i$ and $i+1$ are sequential
  independent, we can use
  Lemma~\ref{lem:indep-global-left}(\ref{lem:indep-global-left:1}) and
  we have that the corresponding steps in positions $i+1$ and $i+2$ in
  $\der{D}_2''$ are still sequential independent. Hence, we can use
  the inductive hypothesis on $\der{D}_2''$ and deduce the existence
  of a switching sequence where $\transp{i+1}[i+2]$ can be switched
  first, i.e. we obtain (see Fig.~\ref{fi:switchB2}, third column)
  \begin{center}
    $\der{D}_2'' \shift{\transp{i+1}[i+2]}
    \der{D}_3''' \shift{\nu_4'''} \ldots \shift{\nu_n'''} \der{D}_n$
  \end{center}

  Putting things together, we obtain
  \begin{center}
    $\der{D}_0 \shift{\transp{i+1}[i+2]} \der{D}_1' \shift{\transp{i}}
    \der{D}_2'' \shift{\transp{i+1}[i+2]} \der{D}_3''' \shift{\nu_4'''}\ldots \shift{\nu_n'''} \der{D}_n$
  \end{center}
  Since we additionally know that in $\der{D}_0$ the steps $i$ and
  $i+1$ are sequentially independent, we can use
  Lemma~\ref{lem:indep-global-left}(\ref{lem:indep-global-left:2}) to
  reorder the first three switchings, obtaining
  \begin{center}
    $\der{D}_0 \shift{\transp{i}} \der{D}_1' \shift{\transp{i+1}[i+2]}
    \der{D}_2'' \shift{\transp{i}} \der{D}_3''' \shift{\nu_4'''}\ldots \shift{\nu_n'''} \der{D}_n$
  \end{center}
  as desired.  The fact that the three switchings above lead to
  $\der{D}_3'''$ is a consequence of Lemma~\ref{lem:switch-confluence}.
\end{proof}
